### 2021-06-08 sections/systematic_uncertainties.tex
first draft version

%Table~\ref{tab:SystSummary} presents a list of all systematic uncertainties considered in the analysis 
%and indicates whether they are taken to be normalisation-only, or to affect both shape and normalisation.
%
%%%%%%%%%%%%%%%
%\begin{table*}[htbp]
%\centering
%\begin{tabular}{lcc}
%\toprule\toprule
%Systematic uncertainty & Type  & Components \\
%\midrule
%\textbf{Luminosity}                  &  N & 1\\\midrule
%\textbf{Reconstructed Objects}                 &   & \\
%Electron                  & SN & 5 \\
%Muon                      &  SN & 6 \\
%Jet reconstruction      & SN & 1\\ 
%Jet vertex fraction         & SN    & 1\\
%Jet energy scale            & SN & 22\\
%Jet energy resolution       & SN & 1\\
%Missing transverse momentum  & SN & 2\\ 
%$b$-tagging efficiency      & SN & 6\\
%$c$-tagging efficiency      & SN & 4\\
%Light-jet tagging efficiency    & SN & 12\\ 
%High-\pt\ tagging   & SN & 1 \\ \midrule
%\textbf{Background Model}                 &   & \\
%$t\bar{t}$ cross section    &  N & 1\\
%$t\bar{t}$ modelling: $\pt$ reweighting   & SN & 9\\
%$t\bar{t}$ modelling: parton shower & SN & 3\\
%$t\bar{t}$+HF: normalisation & N & 2 \\
%$t\bar{t}$+$c\bar{c}$: $\pt$ reweighting  & SN & 2 \\
%$t\bar{t}$+$c\bar{c}$: generator & SN & 4 \\
%$t\bar{t}$+$b\bar{b}$: NLO shape & SN & 8 \\
%$W$+jets normalisation      &  N & 3\\
%$W$ $\pt$ reweighting     &  SN & 1\\
%$Z$+jets normalisation      &  N & 3\\
%$Z$ $\pt$ reweighting     &  SN & 1\\
%Single top normalisation    &  N & 3\\
%Single top model            &  SN & 1\\
%Diboson normalisation  &  N & 3\\
%$t\bar{t}V$ cross section   &  N & 1\\
%$t\bar{t}V$ model           &  SN & 1\\ 
%$t\bar{t}H$ cross section & N & 1 \\
%$t\bar{t}H$ model       & SN & 2 \\ 
%Multijet normalisation  &  N & 4\\ \midrule
%\textbf{Signal Model}                 &   & \\
%$t\bar{t}$ cross section    &  N & 1\\
%Higgs boson branching ratios & N & 3 \\
%$t\bar{t}$ modelling: $\pt$ reweighting   & SN & 9\\
%$t\bar{t}$ modelling: $\pt$ reweighting non-closure  & N & 1\\
%$t\bar{t}$ modelling: parton shower & N & 1\\
%\bottomrule\bottomrule
%\end{tabular}
%\caption{\label{tab:SystSummary} List of systematic uncertainties considered. 
%An ``N'' means that the uncertainty is taken as affecting only the normalisation for all relevant 
%processes and channels, whereas ``SN'' means that the uncertainty is 
%taken on both shape and normalisation.
%Some of the systematic uncertainties are split into several components for a more
%accurate treatment.}
%\end{table*}
%%%%%%%%%%%%%%%
%
%The leading sources of systematic uncertainty vary depending on the analysis channel considered, but they 
%typically originate from $\ttbar$+jets modelling (including $\ttbar$+HF) and $b$-tagging. 
%For example, the total systematic uncertainty  in the background normalisation in the (4 j, 4 b) channel, 
%which dominates the sensitivity in the case of the $\Hc$ search, is approximately 20\%, 
%with the largest contributions originating from $\ttbar$+HF normalisation, $b$-tagging efficiency, 
%$c$-tagging efficiency, light-jet tagging efficiency and $\ttbar$ cross section.
%However, as shown in section~\ref{sec:result}, the fit to data in the nine analysis channels allows the overall background 
%uncertainty to be reduced significantly, to approximately 4.4\%. 
%The reduced uncertainty results from the significant constraints provided by the data on some
%systematic uncertainties, as well as the anti-correlations among sources of systematic uncertainty resulting
%from the fit to the data.
%The total systematic uncertainty in the $\Hc$ signal normalisation in the (4 j, 4 b) channel is approximately 17\%, 
%with similar contributions from uncertainties related to $b$-tagging and overall signal modelling. After the fit,
%this uncertainty is reduced to 7.8\%.
%Table~\ref{tab:SystSummary_WbHc} presents a summary of the sys\-te\-ma\-tic uncertainties for the $\Hc$ 
%search and their impact on the normalisation of the signal and the main backgrounds in the (4 j, 4 b) channel.
%
%The following sections describe each of the systematic uncertainties considered in the analyses. 
%
%\begin{table*}[htbp]
%\centering
%\begin{tabular}{l | c c c c  | c c c c}
%\toprule\toprule
% & \multicolumn{4}{c|}{Pre-fit} & \multicolumn{4}{c}{Post-fit} \\ 
% &  $WbHc$ & $t\bar{t}$+LJ & $t\bar{t}+c\bar{c}$ & $t\bar{t}+b\bar{b}$ &  $WbHc$ & $t\bar{t}$+LJ & $t\bar{t}+c\bar{c}$ & $t\bar{t}+b\bar{b}$ \\
%\midrule
%Luminosity  & $\pm 2.8 $  & $\pm 2.8 $  & $\pm 2.8 $  & $\pm 2.8 $  & $\pm 2.6$ & $\pm 2.6 $  & $\pm 2.6 $  & $\pm 2.6 $ \\ 
%Lepton efficiencies  & $\pm 1.5 $  & $\pm 1.5 $  & $\pm 1.5 $  & $\pm 1.5 $  & $\pm 1.5$ &  $\pm 1.5 $  & $\pm 1.5 $  & $\pm 1.5 $ \\ 
%Jet energy scale  & $\pm 3.3 $  & $\pm 2.9 $  & $\pm 2.3 $  & $\pm 5.8 $  & $\pm 1.4$ &  $\pm 1.2 $  & $\pm 1.8 $  & $\pm 4.1 $ \\ 
%Jet efficiencies  & $\pm 1.2 $  & --  & $\pm 1.9 $  & $\pm 1.7 $  & $\pm 0.9$ &  --  & $\pm 1.4 $  & $\pm 1.2 $ \\
%Jet energy resolution  & --  & $\pm 1.2 $  & $\pm 2.8 $  & $\pm 2.9 $  & -- &  -- & $\pm 1.0 $  & $\pm 1.1 $ \\ 
%$b$-tagging eff.  & $\pm 7.9 $  & $\pm 5.5 $  & $\pm 5.2 $  & $\pm 10 $  & $\pm 5.7$ & $\pm 3.9 $  & $\pm 3.7 $  & $\pm 6.6 $ \\ 
%$c$-tagging eff.  & $\pm 7.0 $  & $\pm 6.6 $  & $\pm 13 $  & $\pm 3.5 $  & $\pm 6.3$ &  $\pm 6.0 $  & $\pm 11 $  & $\pm 3.2 $ \\ 
%Light-jet tagging eff.  & $\pm 0.8 $  & $\pm 18 $  & $\pm 3.2 $  & $\pm 1.5 $  & $\pm 0.6$ &  $\pm 13 $  & $\pm 2.3 $  & $\pm 1.1 $ \\
%$t\bar{t}$: reweighting  & $\pm 5.9 $  & $\pm 2.7 $  & $\pm 4.2 $  & --  & $\pm 3.8$ &  $\pm 1.9 $  & $\pm 2.3 $  & -- \\ 
%$t\bar{t}$: parton shower  & $\pm 5.4 $  & $\pm 4.8 $  & $\pm 10 $  & $\pm 4.9 $  & $\pm 1.7$ &  $\pm 1.5 $  & $\pm 6.5 $  & $\pm 3.1 $ \\ 
%$t\bar{t}$+HF: normalisation  & --  & --  & $\pm 50 $  & $\pm 50 $  & -- &  --  & $\pm 32 $  & $\pm 16 $ \\ 
%$t\bar{t}$+HF: modelling  & --  & --  & --  & $\pm 7.7 $  & -- &  --  & --  & $\pm 7.4 $ \\ 
%Signal modelling  & $\pm 6.9 $  & --  & --  & --  & $\pm 6.9$ &  --  & --  & -- \\ 
%Theor. cross sections  & $\pm 6.2 $  & $\pm 6.2 $  & $\pm 6.2 $  & $\pm 6.2 $  & $\pm 3.9$ &  $\pm 3.9 $  & $\pm 3.9 $  & $\pm 3.9 $ \\ 
%\midrule                                                                                                                                                                                               
%Total   & $\pm 17 $  & $\pm 22 $  & $\pm 54 $  & $\pm 53 $  & $\pm 7.8$ &  $\pm 14 $  & $\pm 28 $  & $\pm 15 $ \\ 
%\bottomrule\bottomrule
%\end{tabular}
%\caption{$\Hcbb$ search: summary of the systematic uncertainties considered in the (4 j, 4 b) 
%channel and their impact (in \%) on the normalisation of the signal and the main backgrounds, before and after the fit to data.
%The $\Hc$ signal and the $t\bar{t}$+light-jets background are denoted by ``$WbHc$''  and ``$t\bar{t}$+LJ'' respectively.
%Only sources of systematic uncertainty resulting in a normalisation change of at least 0.5\% are displayed.
%The total post-fit uncertainty can differ from the sum in quadrature of individual sources due to the 
%anti-correlations between them resulting from the fit to the data.}
%\label{tab:SystSummary_WbHc}
%\end{table*}
%%%%%%%%%%%

@@

The uncertainty in the integrated luminosity is 2.1\%, affecting the overall normalisation of all processes estimated from the simulation. 
%It is derived, following a methodology similar to that detailed in Ref.~\cite{Aaboud:2016hhf}, 
%from a calibration of the luminosity scale using $x$--$y$ beam-separation scans performed in August 2015 and May 2016.
It is derived, following a methodology similar to that detailed in Ref.~\cite{Aaboud:2016hhf}, and using the LUCID-2 detector 
for the baseline luminosity measurements \cite{Avoni:2018iuv}, from a calibration of the luminosity scale using $x$--$y$ beam-separation scans.

@@

%Uncertainties associated with electrons and muons arise from the trigger, reconstruction, identification, and isolation
%efficiencies, as well as the lepton momentum scale and resolution. These are measured in data using 
%$Z\to \ell^+\ell^-$ and $J/\psi\to \ell^+\ell^-$ events~\cite{ATLAS-CONF-2016-024,Aad:2016jkr}. 
%The combined effect of all these uncertainties results in an overall normalisation 
%uncertainty in signal and background of approximately 1\%.

@@

Efficiencies to tag $b$-jets and $c$-jets in the simulation are corrected to match the efficiencies in data by $\pt$-dependent factors,
whereas the light-jet efficiency is scaled by $\pt$- and $\eta$-dependent factors.
The $b$-jet efficiency is measured in a data sample enriched in $\ttbar$ events~\cite{Aaboud:2018xwy}, while the $c$-jet efficiency is measured
using $\ttbar$ events~\cite{ATLAS-CONF-2018-001} or $W$+$c$-jet events~\cite{Aad:2015ydr}. 
The light-jet efficiency is measured in a multijet data sample enriched in light-flavour jets~\cite{ATLAS-CONF-2018-006}.
Since the $\ttbar$ sample used to measure the $c$-jet tagging efficiency overlaps with the analysis sample, the $\Hbb$ search uses
instead the $W$+$c$-jet scale factors.  
In the case of the $\Hbb$ ($\Htautau$) search, the uncertainties in these scale factors include 
a total of 6 independent sources affecting $b$-jets, 1 (2) source(s) affecting $c$-jets, and 17 sources affecting light-jets. 
These systematic uncertainties are taken as uncorrelated between $b$-jets, $c$-jets, and light-jets. 
An additional uncertainty is included due to the extrapolation of these corrections to jets 
with $\pt$ beyond the kinematic reach of the data calibration samples used ($\pt>300~\gev$ for $b$- and $c$-jets, 
and $\pt>750~\gev$ for light-jets); it is taken to be correlated among the three jet flavours. 
%This uncertainty is evaluated in the simulation by comparing the tagging efficiencies while varying e.g.\ the fraction 
%of tracks with shared hits in the silicon detectors or the fraction of fake tracks resulting from random combinations of hits, 
%both of which typically increase at high $\pt$ due to growing track multiplicity and density of hits within the jet. 
Finally, an uncertainty related to the application of $c$-jet scale factors to $\tau$-jets is considered,
but has a negligible impact in these analyses.

@@

A number of sources of systematic uncertainty affecting the modelling of $t\bar{t}$+jets are considered. 
An uncertainty of  6\% is assigned to the inclusive $\ttbar$ production
cross section~\cite{Czakon:2011xx}, including contributions from varying the factorisation and renormalisation 
scales, as well as from the top-quark mass, the PDF and $\alpha_{\textrm{S}}$. The latter two represent the largest contribution 
to the overall theoretical uncertainty in the cross section and were calculated using the PDF4LHC prescription~\cite{Botje:2011sn} 
with the MSTW 2008 68\% CL NNLO, CT10 NNLO~\cite{Lai:2010vv,Gao:2013xoa} and NNPDF2.3 5F FFN~\cite{Ball:2012cx} PDF sets.
The uncertainty associated with the choice of NLO generator is derived by comparing the nominal prediction from
{\powheg}+{\pythiaeight} with a prediction from \textsc{Sherpa}~2.2.1. For the latter, the matrix-element calculation is performed 
for up to two partons at NLO and up to four partons at LO using \textsc{Comix} and \textsc{OpenLoops}, and
merged with the {\sherpa} parton shower using the ME+PS@NLO prescription.
%either {\amcatnlo}+{\pythiaeight} (in the case of the $\Htautau$ search) or
%\textsc{Sherpa} (in the case of the $\Hbb$ search). 
The uncertainty due to the choice of parton shower and hadronisation (PS \& Had) model is derived 
by comparing the predictions from {\powheg} interfaced either to {\pythiaeight} or {\herwig7}. 
The latter uses the MMHT2014 LO~\cite{Harland-Lang:2014zoa} PDF set in combination with the H7UE tune~\cite{Bellm:2015jjp}.
The uncertainty in the modelling of additional radiation is assessed with two alternative {\powheg}+{\pythiaeight} samples:
a sample with increased radiation (referred to as radHi) is obtained by decreasing the renormalisation and factorisation scales  
by a factor of two, doubling the $h_{\textrm{damp}}$ parameter, and using the Var3c upward variation of the A14 parameter set;
a sample with decreased radiation (referred to as radLow) is obtained by increasing the scales by a factor of two 
and using the Var3c downward variation of the A14 set~\cite{ATL-PHYS-PUB-2016-004}.

In the case of the $\Hbb$ search, where the $\ttbar$+HF background plays a prominent role (see Fig.~\ref{fig:Hbb_Summary}), a more 
detailed treatment of its associated systematic uncertainties is used. In particular, since several analysis 
regions have a sufficiently large number of \ttbin\ background events, its normalisation is 
determined in the fit to data.
%, with no prior uncertainty assumed. 
In the case of the \ttcin\ normalisation, an uncertainty of 50\% is assumed, as the fit to the data is unable 
to precisely determine it, and the analysis has very limited sensitivity to this uncertainty.
Since the diagrams that contribute to $\ttbar$+light-jets, \ttcin, and \ttbin\
production are different, all above uncertainties in $\ttbar$+jets
background modelling (NLO generator, PS \& Had, and radHi/radLow), except the uncertainty of the inclusive cross section, are
considered to be uncorrelated among these processes.
Additional uncertainties of the \ttbin\ background are considered associated with the NLO prediction from {\ShOL}, 
which is used for reweighting the nominal {\powheg}+{\pythiaeight} prediction. 
These include three different scale variations, a different shower-recoil model scheme, and 
two alternative PDF sets (MSTW 2008 NLO and NNPDF2.3 NLO). Additional uncertainties are assessed for
the contributions to the \ttbin\ background originating from multiple parton interactions.
Finally, an additional uncertainty is assigned to the \ttbin\ background by comparing
the predictions from {\powheg}+{\pythiaeight} and {\ShOL} 4F (5F vs 4F).
In the derivation of the above uncertainties, the overall normalisations of the \ttcin\ and \ttbin\ backgrounds 
at the particle level are fixed to the nominal prediction. In order to maintain the inclusive $\ttbar$ cross section, 
the normalisation of the $\ttbar$+light-jets background at the particle level is adjusted accordingly.
% Further details are available in Ref.~\cite{Aaboud:2017rss}.

@@

Uncertainties of the data-driven multijet background in the $\Hbb$ search include
contributions from the limited size of the data sample, particularly at high jet and $b$-tag multiplicities, as 
well as from the uncertainty in the rate of fake leptons, estimated in 
different control regions (e.g.\ selected with a upper requirement on either $\met$ or $\mtw$). 
A combined normalisation uncertainty of 50\% due 
to all these effects is assigned, which is taken as correlated across jet
and $b$-tag multiplicity bins, but uncorrelated between electron and muon channels. 
No explicit shape uncertainty is assigned since the large statistical uncertainties associated with
the multijet background prediction, which are uncorrelated 
between bins in the final discriminant distribution, effectively cover all possible shape uncertainties.

Uncertainties of the data-driven fake $\had$ background in the $\Htautau$ search are obtained by using additional signal-depleted regions. The construction is similar to that of the SRs and corresponding CRs discussed in Section~\ref{sec:fakeleptons}, but employing further loosened $\had$ identification criteria, and thus referred to as ``loose SR''  and ``loose CR''. 
%The shape difference between the fake $\tauhad$ background prediction in each loose SR and its associated loose CR
%is assigned as an uncertainty in the prediction in the nominal SR.
%In each loose SR, the shape difference between the data and the fake $\tauhad$ background estimate based on its associated loose CR,
%in both cases having subtracted the small contribution from real $\tauhad$ candidates based on the simulation, 
%is applied as an uncertainty in the prediction in the nominal SR. 
%In each loose SR, the shape difference between the data and the data-driven prediction of its associated loose CR is extracted and applied as an uncertainty in the prediction in the nominal SR. 
In each loose SR, after subtracting the small simulation-predicted contribution from real $\had$ candidates, 
the relative difference in the shape of the distribution between the remaining data and the fake $\had$ background estimate based on its associated loose CR 
is assigned as an uncertainty of the prediction in the nominal SR.
In addition, a 30\% uncertainty is applied to the fraction of $\ttbar$ events with a fake $\had$ candidate from the simulation that are added to the fake $\had$ template in the $\hadhad$ channel as part of the fake $\had$ background estimation procedure. This uncertainty, associated with the modelling of the fake $\had$ rate by the simulation, is estimated by comparing data and simulation in a sample enriched in $\ttbar$ dilepton events plus a fake $\had$ candidate. 
The same uncertainty is assigned to the selected signal events with fake $\had$ candidates.
In addition, a systematic uncertainty is assigned to account for the different fractional composition of particles (various types of leptons and partons) producing the fake $\had$ candidates 
between each SR and its corresponding CR in the $\ttbar$ simulation. 
Finally, the normalisation of the fake $\had$ background in each SR is determined in the fit to data. 
%In order to be conservative, the yield of different origins of the fake taus is studied and a dedicated uncertainty is added to the fake template. The MC studies shows that non-H-to-tautau signal with fake taus can contribute to the $\lephad$ channel. So a 30\% uncertainty is  applied to these signal-faking-tau events.

@@

Several normalisation and shape uncertainties are taken into account for the $\Hq$ signal.
The uncertainty of the $\ttbar$ cross section also applies to the $\Hq$ signal and is taken to be the same as, 
and fully correlated with, the uncertainty assigned to the $\ttbar \to WbWb$ background.
Uncertainties of the Higgs boson branching ratios are taken into account
following the recommendation in Ref.~\cite{deFlorian:2016spz}.
%: $\pm 1.1\%$ ($\Delta\alpha_\textrm{S}$),  $\pm 1.4\%$ ($\Delta m_b$) and $\pm 0.8\%$ (theory).
Additional uncertainties associated with the modelling of additional radiation, with the choice of NLO generator, and
with the choice of parton shower and hadronisation model, are estimated from the comparison of the nominal
and alternative $\ttbar \to WbWb$ background samples (discussed in Section~\ref{sec:syst_bkgmodeling}) and applied to $\Hq$ signal. 
These modelling uncertainties are taken to be uncorrelated with those affecting the $\ttbar \to WbWb$ background.

>>>>

@@

The uncertainty in the integrated luminosity is 1.7\%, affecting the overall normalisation of all processes estimated from the simulation. 
It is derived, following a methodology similar to that detailed in Ref.~\cite{Aaboud:2016hhf}, and using the LUCID-2 detector 
for the baseline luminosity measurements \cite{Avoni:2018iuv}, from a calibration of the luminosity scale using $x$--$y$ beam-separation scans.

@@

@@

Efficiencies to tag $b$-jets and $c$-jets in the simulation are corrected to match the efficiencies in data by $\pt$-dependent factors,
whereas the light-jet efficiency is scaled by $\pt$- and $\eta$-dependent factors.
The $b$-jet efficiency is measured in a data sample enriched in $\ttbar$ events~\cite{Aaboud:2018xwy}, while the $c$-jet efficiency is measured
using $\ttbar$ events~\cite{ATLAS-CONF-2018-001} or $W$+$c$-jet events~\cite{Aad:2015ydr}. 
The light-jet efficiency is measured in a multijet data sample enriched in light-flavour jets~\cite{ATLAS-CONF-2018-006}.
The uncertainties in these scale factors include a total of 44 independent sources affecting $b$-jets, 19 source affecting $c$-jets, and 19 sources affecting light-jets. 
These systematic uncertainties are taken as uncorrelated between $b$-jets, $c$-jets, and light-jets.

@@

A number of sources of systematic uncertainty affecting the modelling of $t\bar{t}$+jets are considered. 
An uncertainty of  6\% is assigned to the inclusive $\ttbar$ production
cross section~\cite{Czakon:2011xx}, including contributions from varying the factorisation and renormalisation 
scales, as well as from the top-quark mass, the PDF and $\alpha_{\textrm{S}}$. The latter two represent the largest contribution 
to the overall theoretical uncertainty in the cross section and were calculated using the PDF4LHC prescription~\cite{Botje:2011sn} 
with the MSTW 2008 68\% CL NNLO, CT10 NNLO~\cite{Lai:2010vv,Gao:2013xoa} and NNPDF2.3 5F FFN~\cite{Ball:2012cx} PDF sets.
The uncertainty associated with the choice of NLO generator is derived by comparing the nominal prediction from
{\powheg}+{\pythiaeight} with a prediction from \textsc{Sherpa}~2.2.1. For the latter, the matrix-element calculation is performed 
for up to two partons at NLO and up to four partons at LO using \textsc{Comix} and \textsc{OpenLoops}, and
merged with the {\sherpa} parton shower using the ME+PS@NLO prescription.

The uncertainty due to the choice of parton shower and hadronisation (PS \& Had) model is derived 
by comparing the predictions from {\powheg} interfaced either to {\pythiaeight} or {\herwig7}. 
The latter uses the MMHT2014 LO~\cite{Harland-Lang:2014zoa} PDF set in combination with the H7UE tune~\cite{Bellm:2015jjp}.
The uncertainty in the modelling of additional radiation is assessed with two alternative {\powheg}+{\pythiaeight} samples:
a sample with increased radiation (referred to as radHi) is obtained by decreasing the renormalisation and factorisation scales  
by a factor of two, doubling the $h_{\textrm{damp}}$ parameter, and using the Var3c upward variation of the A14 parameter set;
a sample with decreased radiation (referred to as radLow) is obtained by increasing the scales by a factor of two 
and using the Var3c downward variation of the A14 set~\cite{ATL-PHYS-PUB-2016-004}.

@@

The statistical uncertainties of the fake $\had$ background calibration in the leptonic channel is applied with uncorrelated uncertainties for different fake $\had$ source and different $\pt$ slices. The uncertainties of the ABCD method is applied with a normalisation factor for muon and electron respectively considering both statistical fluctuation and the differences between signal regions.

The uncertainties of the fake factor method applied in hadronic channel includes statistical uncertainties of each fake factor and different sets of fake factors derived from different signal depleted CRs.

@@

Several normalisation and shape uncertainties are taken into account for the $\Hq$ signal.
Uncertainties of the Higgs boson branching ratios are taken into account
following the recommendation in Ref.~\cite{deFlorian:2016spz}.
The uncertainty of ISR, FSR and PDF are treated correlatedly with $\ttbar$ samples. The parton shower uncertainties is estimated by comparing the predictions from {\powheg} interfaced either to {\pythiaeight} or {\herwig7}.
